Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-03-17 VegOzone.tex
Change reference to bibtex file
Write about results

We have modeled $\mathrm{POD_y}$ (flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA)) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variaty of bespoke parameterizations in $f_\mathrm{T}$, $f_\mathrm{light}$, and GS (described in detail in Section~\ref{subsec:do3se_parameters}). The results are shown comprehensively in Fig.~\ref{fig:pody_rel}. For comparison, critical levels (CLs) for a $4\,\unit{\%}$ ($2\,\unit{\%}$ for Norway spruce) reduction in biomass in accordance to \citet{ICP:MappingManual2017} are shown as horizontal dashed lines. Different years are color coded. Lighter hue indicates simulations with MM GS and MM canopy and leaf parameters. Annotated $\pm$ refers to bespoke light response functions, while bespoke temperature response functions are indicated by categories on the x-axis. Closed/open symbols represent SWP sensitivity switched off/on in the model.

For downy birch (Fig.~\ref{fig:pody_rel}a), all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and display the largest spread among the species ($16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 for boreal type). As expected, SWP reduces ozone uptake in 2018 mainly. The uncertainty induced by the light sensitivity threshold ranges between $(1.5..2.5)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $(-2.2..-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. With MM GS, 2018 displays increasingly larger ozone uptake than 2019 with more acclimated paramters. However accounting for the shorter and later GS in 2018, we find a reversal of this relationship in the boreal type. We deduce an overall systematic uncertainty of $xx$.

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce displays an overall lower spread ($10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 for boreal type) than downy birch, but appears more sensitive to GS and interannual veriability in conditions (boreal GS MM compared to GS bespoke). The sensitivity to the light threshold is also smaller compared to downy birch. Accounting for the shorter and later GS in 2018, we find again a reversal between predicted ozone damage in 2018 and 2019 for the boreal type. We deduce an overall systematic uncertainty of $xx$.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. CL is only exceeded for boreal types and cold type in 2018. We find the same reveral between 2018 and 2019 as for the other species in the boreal type. Perannial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant for the MM GS. The spread in 2018 for boreal type is of the order of $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. We deduce an overall systematic uncertainty of $xx$.

These results allow for an assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake:
\begin{itemize}
\item Optimal acclimation of species or PFT $\rightarrow$ increased uptake in any year,
\item Identification of proper timing of the growing season $\rightarrow$ reduced uptake in any year,
\item Drought conditions $\rightarrow$ reduced uptake,
\item Early onset of growing season $\rightarrow$ increased uptake.
\end{itemize}
The exact order of magnitude of these effects varies with species or PFT and from year to year. The systematic uncertainty induced by our proposed temperature acclimation is, in general, larger than for a $\pm 20\,\unit{\%}$ variation in light response threshold and can likely exceed $xx\,\unit{\%}$ for some years and species. This would lead to a huge underestimation of induced ozone damage. At the same time, the magnitude of light response uncertainty is increasing with acclimation to colder temperature. The impact of the timing of the GS can be at the same oder of magnitude as the acclimation of response function but points into the oposite direction. Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak (compare 2018 GS MM and GS bespoke) induces large uncertainties which are increasing if an acclimation to colder climate is anticipated.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig11}
  \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd. Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhoved (2018, 2019). (a) Downy birch; (b) Norway spruce; (c) Perennial grassland. The dashed horizontal lines indicate the critical levels defined in \citet{ICP:MappingManual2017} for each species, respectively.}
  \label{fig:pody_rel}
\end{figure}

@@

@@

Manninen et al. (2009a) summarized the results from open-top chamber (OTC) experiments performed in northern Finland where Scots pine and northern birch provenances were exposed to 36 – 54 ppb O3 as mean daylight concentrations. Biomass reductions of both downy birch (Betula pubescens) and mountain birch seedlings were demonstrated after one-season exposure. Moreover, the root:shoot ratio was reduced and the leaves of the mountain birch in the high O3 treatments showed necrotic stipples already in June. Mountain birch has a very rapid phenological development during spring-early summer that makes it difficult for this species to compensate for damaged leaves. It is also noteworthy that N fertilization increased the total biomass production of mountain birch markedly only under the ambient O3 both under the current climate (open-field plots) and in the about 1°C warmer climate (non-filtered air OTCs) but not under elevated O3 (Manninen et al. 2009b).
In contrast to mountain birches, northern Scots pine of local origin only showed invisible needle-level O3 impacts, such as increased respiration, decreased level of polyamines and reduced chlorophyll/carotenoid ratio and increased amount of epicuticular waxes after three years of exposure (Manninen et al. 2009a). There is no field data comparing the O3 sensitivity of birches and Scots pine from northern Europe, but both Scots pine and Norway spruce have been ranked as more O3-tolerant species than deciduous birches and Alnus incana based on crown condition in a Lithuanian study (Girgzdiene et al. 2009). Despite the relatively high O3 tolerance of Scots pine, invisible needle injuries may not be excluded in early spring conditions with high O3 concentrations though (Augustaitis and Bytnerowicz 2008, Huttunen and Manninen 2013).

@@

\begin{figure*}[t]
  \includegraphics[width=12cm]{figA2}
  \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia. Jergul/Karasjok is well correlated with Esrange, and Pallas. Svanvik displays highest correlation with Pallas. (a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
  \label{fig:density_distribution}
\end{figure*}

@@

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_birch}\\
  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
  \includegraphics[width=12cm]{figB3}
\caption{Construction of bespoke response functions for Norway birch. (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively. Original mapping manual parameterization is shown in comparison as solid line. Note that $Q_0$ has been truncated to $0.006$. PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
\label{fig:f_temp_birch}
\end{figure}

\begin{figure*}[t]
  \includegraphics[width=12cm]{figB4}
  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (\unit{doy}), March--October. A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen. \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$. (a, b) Birch; (c, d) Norway spruce; (e, f) perennial grassland. From left to right: 2018, 2019. {\bf TODO: substitute with figure for Boreal PFTs?}}
  \label{fig:pody_mm_composit}
\end{figure*}

@@

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
\bibliographystyle{copernicus}
\bibliography{VegOzone.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}

>>>>

We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variaty of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}). The results are shown comprehensively in Fig.~\ref{fig:pody_rel}. For comparison, critical levels for a $4\,\unit{\%}$ (silver birch), $2\,\unit{\%}$ (Norway spruce), and $10\,\unit{\%}$ (perennial grassland, above ground) reduction in biomass \citep{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. Different years are color coded as previously. Simulations with MM GS, canopy, and leaf parameters are indicated with squared markers, bespoke GS with circles. The annotation $\pm$ refers to bespoke light response function threshold ($+$ increased, $-$ decreased), while bespoke temperature response functions are indicated by categories on the x-axis. Closed/open symbols represent SWP sensitivity switched off/on in the model.

The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake with respect to MM default parameterization:
\begin{itemize}
  \itemsep0pt
\item Increasing temperature acclimation $\rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed onset of growing season $\rightarrow$ increased uptake,
\item drought conditions (SWP) $\rightarrow$ reduced uptake,
\item lower/higher light threshold $\rightarrow$ incresed/reduced uptake,
\item bespoke light threshold $\rightarrow$ asymmetric response in POD,
\item temperature acclimation $\rightarrow$ amplified effect of light threshold and drought.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species. Drought effects (SWP) do only matter if a bespoke GS is not taken into account.

In more detail for downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Downy birch displays the largest range of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light threshold is especially pronounced, e.g. $\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2019. Without bespoke GS 2018 stands out as a higher ozone damage risk year compared to 2019. However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type. SWP induced reduction in ozone uptake is only effectiv in 2018 and without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce is the least sensitive to acclimation in temperature and therefor the least biased. Norway spruce displayes a medium sensitivity to the light threshold, e.g. $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018. Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. Perennial grassland shows the smallest ozone uptake, but a similarily large bias as downy birch. We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type. Perannial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig11}
  \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd. Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhoved (2018, 2019). (a) Downy birch; (b) Norway spruce; (c) Perennial grassland. The dashed horizontal lines indicate critical levels defined in \citet{ICP:MappingManual2017,ESPR:Hayes2021} for each species, respectively.}
  \label{fig:pody_rel}
\end{figure}

@@

---
The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake with respect to MM default parameterization:
\begin{itemize}
\item Increasing temperature acclimation $\rightarrow$ increased uptake regardless of individual year,
\item Usage of proper growing season $\rightarrow$ reduced uptake with respect to the above,
\item An earlier assumed onset of growing season $\rightarrow$ increased uptake,
\item Drought conditions (SWP) $\rightarrow$ reduced uptake,
\item Lower/higher light threshold $\rightarrow$ incresed/reduced uptake.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species.
 The impact of the timing of the GS can be at the same oder of magnitude as the acclimation of response function but points into the oposite direction. Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak induces large uncertainties in anticipation of an acclimation to colder climates.

@@

Manninen et al. (2009a) summarized the results from open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch provenances were exposed to 36 – 54 ppb O3 as mean daylight concentrations. Biomass reductions of both downy birch (Betula pubescens) and mountain birch seedlings were demonstrated after one-season exposure. Moreover, the root:shoot ratio was reduced and the leaves of the mountain birch in the high O3 treatments showed necrotic stipples already in June. Mountain birch has a very rapid phenological development during spring-early summer that makes it difficult for this species to compensate for damaged leaves. It is also noteworthy that N fertilization increased the total biomass production of mountain birch markedly only under the ambient O3 both under the current climate (open-field plots) and in the about 1°C warmer climate (non-filtered air OTCs) but not under elevated O3 (Manninen et al. 2009b).
In contrast to mountain birches, northern Scots pine of local origin only showed invisible needle-level O3 impacts, such as increased respiration, decreased level of polyamines and reduced chlorophyll/carotenoid ratio and increased amount of epicuticular waxes after three years of exposure (Manninen et al. 2009a). There is no field data comparing the O3 sensitivity of birches and Scots pine from northern Europe, but both Scots pine and Norway spruce have been ranked as more O3-tolerant species than deciduous birches and Alnus incana based on crown condition in a Lithuanian study (Girgzdiene et al. 2009). Despite the relatively high O3 tolerance of Scots pine, invisible needle injuries may not be excluded in early spring conditions with high O3 concentrations though (Augustaitis and Bytnerowicz 2008, Huttunen and Manninen 2013).

@@

\begin{figure*}[t]
  \includegraphics[width=12cm]{figA2}
  \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia. Jergul/Karasjok is well correlated with Esrange, and Pallas. Svanvik displays highest correlation with Pallas. Note: The stripe at $26\,\unit{ppb}$ is an artifact from plotting. (a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
  \label{fig:density_distribution}
\end{figure*}

@@

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_birch}\\
  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
  \includegraphics[width=12cm]{figB3}
\caption{Construction of bespoke response functions for downy birch. (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively. Original mapping manual parameterization is shown in comparison as solid line. Note that $Q_0$ has been truncated to $0.006$. PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
\label{fig:f_temp_birch}
\end{figure}

\begin{figure*}[t]
  \includegraphics[width=12cm]{figB4}
  \caption{$\mathrm{DO_3SE}$ modeling results for mapping manual default parameterazation. $\mathrm{POD_y}$ is shown over day of year (\unit{doy}), March--October. A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen. \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$. (a, b) Downy birch; (c, d) Norway spruce; (e, f) perennial grassland. From left to right: 2018, 2019.}
  \label{fig:pody_mm_composit}
\end{figure*}

@@

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
\bibliographystyle{copernicus}
\bibliography{atmo.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}
